---
title: "Homework 2"
output: github_document
---

Here I am using the `mpg` dataset to investigate how gas mileage varies among cars of different classes.

### Load Tidyverse
```{r setup, message=FALSE}

# If it's the first time you've ever used the package:
# install.packages("tidyverse")

# Load the tidyverse
library(tidyverse)
```

### Inspect data
```{r explore_the_dataframe}

glimpse(mpg) # A tidyverse way to inspect the tibble



```

### Visualize data
Here I plot mileage against vehicle class. I also consider the number of cylinders, since this a variable that affects mileage and varies among vehicle classes.
```{r, message=FALSE}
ggplot (data = mpg) +
  geom_jitter(mapping = aes(x=class, y=cty, color = as.character(cyl)), alpha=.8, size=2, width = .2) +
  scale_color_ordinal() +
  theme_dark() +
  theme(panel.grid.major.x = element_blank()) +
  labs(
    color = "Number of cylinders",
    x = "Vehicle class",
    y = "Miles per gallon (city)"
  )


```

This plot indicates that there are distinct patterns in the mileage of different vehicle classes, though there is overlap among all classes. However, from looking at this plot I would probably conclude that the number of  cylinders is a more important determinant of gas mileage than vehicle class.


```{r, include=FALSE}

ggplot (data = mpg) +
  geom_jitter(mapping = aes(x=as.character(cyl), y=displ, color = cty), size=2.5, width = .18) +
  scale_color_viridis_c() +
  theme_light() +
  labs(
    color = "Miles per gallon (city)",
    x = "Number of cylinders",
    y = "Engine displacement (liters)"
  )

```

