---
title: "Homework 5"
output: github_document
---
### Setup
```{r, message=FALSE}
# Load the tidyverse
library(tidyverse) 

# Look at the data
glimpse(diamonds)
```
  
### 1. Print a data frame that contains the simple count of diamonds in each cut
```{r}
# count(group_by(diamonds,cut))
# piped:
cut_summary <- diamonds %>%
                group_by(cut) %>%
                  count()
cut_summary
```

  
### 2. Use dplyr to create a df that contains 1% of the rows of `diamonds`, randomly subsetted
```{r}
random_subset <- diamonds %>%
                  slice_sample(prop = 0.01) 
# slices out a random selection of rows, 0.01 the size of the original data frame

glimpse(random_subset)

```

  
### 3. Calculate the average size of the 100 largest diamonds in each clarity category
```{r}
largest.by.clarity <- diamonds %>%
                      group_by(clarity) %>%
                      slice_max(carat, n = 100, with_ties = FALSE) %>%  # takes the 100 top values by carat
                      summarise(mean.size = mean(carat, na.rm = TRUE))  # makes a summary df of average carats

largest.by.clarity
```
  
### 4. Make scatterplots of dimensions
```{r}
# First, a plot of x vs. y
ggplot(diamonds) +
  geom_point(mapping = aes(x = x, y = y)) +
  labs(x = "Length (mm)", y = "Width (mm)")

# Next, a plot of x vs. z
ggplot(diamonds) +
  geom_point(mapping = aes(x = x, y = z)) +
  labs(x = "Length (mm)", y = "Depth (mm)")

```
  
### 5. Create a new data frame, with these questionable points removed
```{r}
sans_outliers <- diamonds %>%
                  filter(x > 3 & y < 50 & z < 30) # filtering out the most extreme outliers in the plots above
```
I left some less extreme outliers because I dunno what's normal for diamonds. Maybe they can be long and skinny like a hotdog? I doubt it but I do hope so