---
title: "Homework 7: BAD PLOT!"
output: github_document
---

## Bad plot
```{r bad_plot, message=FALSE, warning=FALSE}

library(tidyverse)
library(ggmap)

## Import data
data <- read.csv("hw_7_data.csv")
glimpse(data)

## Get map
bad.map <- good.map <- get_map(center=c(-122.6, 37.9), source = "stamen", maptype = "watercolor", zoom =10)

## Plot data
ggmap(bad.map)  +
  geom_point(data = data, mapping = aes(x = jitter(WEST.COORD, amount = .1), y = jitter(NORTH.COOR, amount = .06), color = Laeq, fill = trill.bw, shape = Type, size = Dialect), stroke = 3) +
  scale_shape_manual(values = c(21, 22)) +
  scale_colour_distiller(palette = "YlOrRd", direction = 1) +
  scale_fill_distiller(palette = "BuGn", direction = 1)

```

#### Analysis
The gimmicky watercolor map was *probably* not the best choice. It lacks detail and the colors are garish. 

There are too many aesthetic mappings for the map data to be at all comprehensible; not only is point position mapped to latitude and longitude, but each point also has a **shape** mapping (habitat type), **size** mapping (song dialect), and not one but **two color** mappings -- an inner fill color mapped to trill bandwidth, and an outline color mapped to background noise. 

Though it would be acceptable to map either trill bandwidth or background noise to a continuous color scale, the color combination is difficult to interpret (and to look at). The colors chosen for these two color scales are not only as unfriendly as possible to colorblind viewers, but also unhelpfully match the over-the-top map colors. 

The shape mapping is pointless, as the habitat types do not overlap anywhere on the map. 

The size mapping is an absolutely terrible choice to distinguish among song dialects, which are unordered factor levels. Not only does the size scale give a false impression of both ordered and continuous data, but it's also impossible to distinguish most of the sizes. The large size of many of the points also exacerbates the bad overplotting.

The greatest sin in this horrible graphic is the way it addresses overplotting: it jitters the points. While slight jitter may be appropriate in a scatterplot, datapoints on a map are supposed to accord with an actual geographical position and jittering them is a dishonest representation of the data. Moreover, good jitter is supposed to happen on a small enough scale that it doesn't alter the interpretation of the data. Here the points are jittered by as much as 0.1 degrees -- the equivalent of nearly seven miles. This makes it look as though birds were sampled continuously across the Bay Area (including, apparently, in the Pacific Ocean), when in reality they were sampled from 6 discrete study sites.







## Marginally better plot

The premise of this plot isn't great, as it's probably too much data to include in one figure, but let's try to fix it.

Unjittered, the points wouldn't show up clearly at this scale, since they're concentrated in 6 small study sites. I'll therefore facet the map into 6 more zoomed-in maps of these study sites. Of course, I will change the base map to something less gaudy and more informative. I'll pull out the aesthetic mappings for habitat type (urban/rural), song dialect, and background noise and instead include it as a text label on the faceted grid.





```{r marginally_better_plot, message=FALSE, warning=FALSE}
## Setup
library(tidyverse)
library(ggmap)
library(cowplot)

## Create a summary data frame -- mean background noise for each dialect
dialect.summary <- data %>%
  group_by(Dialect) %>%
  summarize(mean = mean(Laeq), lon = mean(WEST.COORD), lat = mean(NORTH.COOR))
glimpse (dialect.summary)

## Make maps! One for each dialect. Would be easier in qGIS
presidio <- ggmap(get_map(center=c(-122.4751, 37.80025), source = "stamen", maptype = "terrain", zoom = 15))
lame <- ggmap(get_map(center=c(-122.5015, 37.71755), source = "stamen", maptype = "terrain", zoom = 15))
richmond <- ggmap(get_map(center=c(-122.3435, 37.90985), source = "stamen", maptype = "terrain", zoom = 15))
commonweal <- ggmap(get_map(center=c(-122.7240, 37.91065), source = "stamen", maptype = "terrain", zoom = 15))
limantour <- ggmap(get_map(center=c(-122.8828, 38.02837), source = "stamen", maptype = "terrain", zoom = 15))
abla  <- ggmap(get_map(center=c(-122.9448, 38.12043), source = "stamen", maptype = "terrain", zoom = 15))


SF <- presidio +
  geom_point(data = data, mapping = aes(x = WEST.COORD, y = NORTH.COOR, color = trill.bw), size = 2.5) +
  scale_color_distiller(palette = "RdPu") +
  theme(axis.ticks = element_blank(), axis.text = element_blank(), axis.title.x = element_blank(), legend.position = "none") +
  labs(title = "San Francisco", subtitle = "Mean noise (dB) = 56.00", y = "URBAN")

SF.LM <- lame +
  geom_point(data = data, mapping = aes(x = WEST.COORD, y = NORTH.COOR, color = trill.bw), size = 2.5) +
  scale_color_distiller(palette = "RdPu") +
  theme(axis.ticks = element_blank(), axis.text = element_blank(), axis.title.x = element_blank(), legend.position = "none") +
  labs(title = "Lake Merced", subtitle = "Mean noise (dB) = 55.16", y = "")

berkeley <- richmond +
  geom_point(data = data, mapping = aes(x = WEST.COORD, y = NORTH.COOR, color = trill.bw), size = 2.5) +
  scale_color_distiller(palette = "RdPu") +
  theme(axis.ticks = element_blank(), axis.text = element_blank(), axis.title.x = element_blank(), legend.position = "none") +
  labs(title = "Berkeley", subtitle = "Mean noise (dB) = 51.29", y = "")

clear <- commonweal +
  geom_point(data = data, mapping = aes(x = WEST.COORD, y = NORTH.COOR, color = trill.bw), size = 2.5) +
  scale_color_distiller(palette = "RdPu") +
  theme(axis.ticks = element_blank(), axis.text = element_blank(), axis.title.x = element_blank(), legend.position = "none") +
  labs(title = "Clear", subtitle = "Mean noise (dB) = 48.20", y = "RURAL")

lima <- limantour +
  geom_point(data = data, mapping = aes(x = WEST.COORD, y = NORTH.COOR, color = trill.bw), size = 2.5) +
  scale_color_distiller(palette = "RdPu") +
  theme(axis.ticks = element_blank(), axis.text = element_blank(), axis.title.x = element_blank(), legend.position = "none") +
  labs(title = "Limantour", subtitle = "Mean noise (dB) = 47.57", y = "")

mcclure <- abla +
  geom_point(data = data, mapping = aes(x = WEST.COORD, y = NORTH.COOR, color = trill.bw), size = 2.5) +
  scale_color_distiller(palette = "RdPu") +
  theme(axis.ticks = element_blank(), axis.text = element_blank(), axis.title.x = element_blank(), legend.position = "none") +
  labs(title = "McClure", subtitle = "Mean noise (dB) = 43.53", y = "")


## A dummy plot to extract the legend from...
dummyplot <- ggplot(data, aes(x = WEST.COORD, y = NORTH.COOR, color = trill.bw)) +
  geom_point() +
  scale_color_distiller(palette = "RdPu", name = str_wrap("Trill bandwidth (Hz)", width = 3))


## Can't facet wrap this, so I'm arranging with ggdraw and the package cowplot
toprow <- plot_grid(SF, SF.LM, berkeley, align = "h", axis = "bt", ncol = 3)
bottomrow <- plot_grid(clear, lima, mcclure, align = "h", axis = "bt", ncol = 3)

legend <- get_legend(dummyplot)

ggdraw(plot_grid(plot_grid(toprow, bottomrow, ncol=1, align='v'),
                 plot_grid(NULL, legend, ncol=1),
                 rel_widths=c(1, .15), greedy = FALSE))
```