---
title: "Homework 8"
output: github_document
---

## Question 1
Using the `diamonds` data set, make a data frame that shows the slope of the relationship of price as a function of weight (carat) for each unique color, in decreasing order of slope.

``` {r setup_Q1, message=FALSE, echo=FALSE}
library(tidyverse)
data("diamonds")
```


Visualizing as a first step, to conceptualize what I'm supposed to do
``` {r visualize_Q1, message = FALSE}
 ggplot(diamonds) +
  geom_smooth(aes(x = carat,
                  y = price,
                  color = color),
              method = lm)
```

Make the data frame:
``` {r Q1, warning = FALSE}
# Nest data (each row is a group)
by_color <- diamonds %>%
  group_by(color) %>%
  nest()

# Make a function that runs the model I want
price.carat.model <- function(df) {
  lm(price ~ carat, data = df)
}

# Apply the function to every group in my nested data frame
models <- map(by_color$data, price.carat.model)

# Add the list of models to my data frame
by_color <- by_color %>%
  mutate(model = map(data, price.carat.model))
by_color

# Expand the model contents into columns
model_by_color <- by_color %>%
  mutate(tidy = map(model, broom::tidy)) %>%
  unnest(tidy, .drop = TRUE) %>%
  filter(term == "carat")
model_by_color

# Create final data frame
model_by_color %>%
  summarise(price.per.carat = estimate) %>%
  arrange(desc(price.per.carat))
```


## Question 2: nonlinear least squares models
### Fitting nonlinear models


```{r setup_Q2}
library(nls2) ## "Non-linear regression with brute force"
data("DNase") ## ELISA band optical density as a function of protein concentration, for 12 different runs
glimpse(DNase)
```

Visualize the data
``` {r visualize_Q2}
ggplot(DNase) +
  geom_point(aes(x = conc,
                 y = density))
```

Use list-columns to create a separate square root model and Monod model for density as a function of concentration, for each run. Extract AIC scores to evaluate model quality.
```{r Q2}
# Nest data
by_run <- DNase %>%
  group_by(Run) %>%
  nest()

# Write functions for the two models
sqrt.mod <- function(df) {
  nls2(formula(density ~ beta_1 * sqrt(conc) + beta_0),
        data = df,
        start = list(beta_1 = 0.5, beta_0 = 0.1))
}

monod.mod <- function(df) {
  nls2(formula(density ~ (conc * beta_1)/(conc + beta_0)),
       data = df,
       start = list(beta_1 = 2, beta_0 = 0.5))
}


# Apply my two models to every group in my data frame

sqrt_models <- map(by_run$data, sqrt.mod)
monod_models <- map(by_run$data, sqrt.mod)

# Add the lists of models to my data frame

by_run <- by_run %>%
  mutate(sqrt.model = map(data, sqrt.mod), 
         monod.model = map(data, monod.mod))
by_run

# Expand model contents into columns
## Use broom::glance to extract AIC 

models_by_run <- by_run %>%
  mutate(glance.sqrt = map(sqrt.model, broom::glance),
         glance.monod = map(monod.model, broom::glance)) %>%
  unnest(c(glance.sqrt, glance.monod), .drop = TRUE)



summary <- models_by_run %>%
  select(Run, AIC, AIC1)

```

Make a plot of AIC scores for each model for each run
```{r visualization}

ggplot(summary) +
  geom_point(aes(x = as.integer(Run), y = AIC), size = 2.5, color = "red") +
  geom_point(aes(x = as.integer(Run), y = AIC1), size = 2.5) +
  annotate("text", x = 4, y = -38, label = "Square-root model", color = "red") +
  annotate("text", x = 4, y = -63, label = "Monod model") +
  scale_x_continuous(name = "Run", breaks = c(1:11)) +
  theme_minimal()
```

Since the Monod model had lower AIC values than the square-root model across all the runs, it's probably a better fit for these data.

## Extra credit

```{r predictions}

library(modelr)

m1 <- nls2(formula(density ~ beta_1 * sqrt(conc) + beta_0),
        data = DNase,
        start = list(beta_1 = 0.5, beta_0 = 0.1))

m2 <- nls2(formula(density ~ (conc * beta_1)/(conc + beta_0)),
       data = DNase,
       start = list(beta_1 = 2, beta_0 = 0.5))
        
        
DNase_pred <- DNase %>%
  gather_predictions(m1, m2)


ggplot(data = DNase_pred) +
  geom_point(aes(x = conc, y = density)) +
  geom_smooth(aes(x = conc, y = density), size = 2, color = "black") +
  geom_smooth(aes(x = conc, y = pred, color = model), linetype = "dashed") +
  scale_color_discrete(labels = c("Square-root model","Monod model")) +
  labs(color = "Predicted values",
       x = "Protein concentration",
       y = "Optical density")
```